---
output: github_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)

View(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

Similarly, deal with the returns of characters.

Based on these datasets calculate the average number of deaths an Avenger suffers. 

```{r}
# Deaths

deaths <- av %>% 
  pivot_longer(
    starts_with("Death"),
    names_to = "Time",
    values_to = "Died"
  ) %>% 
  select(
    URL,
    Time, Died
  ) %>% 
  mutate(
    Time = parse_number(Time)
    ) %>%
  group_by(URL,Died) %>%
  summarise(deaths = max(Time)) %>%
  filter(URL != "") %>%
  filter(Died != "", Died != "NO")


# Calculate the average number of deaths an Avenger suffers
# Filter out blank values and count "yes" entries
avg_deaths <- deaths %>%
  group_by(URL) %>%
  summarise(death_count = n()) %>%
  summarise(average_deaths = mean(death_count))

  
avg_deaths


# returns 
returns <- av %>% 
  pivot_longer(
    starts_with("Return"),
    names_to = "Time",
    values_to = "Return"
  ) %>% 
  select(
    URL,
    Time, Return
  ) %>% 
  mutate(
    Time = parse_number(Time)
    ) %>%
  group_by(URL,Return) %>%
  summarise(returns = max(Time)) %>%
  filter(URL != "") %>%
  filter(Return != "")
  
View(returns)



```

## Individually

For each team member, copy this part of the report. 

#Alister
Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

```{r}

# Calculate the number of unique Avengers who have died at least once
unique_deaths <- deaths %>%
  filter(deaths >= 1, Died == "YES") %>%
  summarise(died_at_least_once = n_distinct(URL))

# Calculate the total number of Avengers from the original dataset
total_avengers <- av %>%
  summarise(total = n_distinct(URL))

# Display the results
cat("Total Avengers:", total_avengers$total, "\n")
cat("Avengers who died at least once:", sum(unique_deaths$died_at_least_once), "\n")





```



### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

#Brietta
Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> 57 occasions the individual made a comeback.

### Code

```{r}
return_count = returns %>%
  filter(returns >= 2)%>%
  summarise(return_count = n()) %>%
  summarise(return_count = sum(return_count)) %>%
  ungroup()


return_count
```

### Answer

My code shows that only 17 times the avengers had a return after dying, which is much lower than the articles claim of 57 ocasions that someone came back


#Brianna
Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.
